Methods for Analyzing Models of Rod-Shaped Bacteria

### 2025-07-23 paper/main.tex
add figure for final and initial states
remove section on data generation
move orcid.pdf to figures folder
create standalone file with overview

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[hidelinks]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{float}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{mathrsfs}
\usepackage{onimage}
\usepackage{graphicx}
\usepackage{doi}
\usepackage{acronym}
\usepackage{listings}
\usepackage{tikz}
\usepackage{tabularx}
\usepackage{siunitx}
\usepackage{cite}
\usepackage{acronym}
\usepackage{xcolor}

\usetikzlibrary{trees}
\usetikzlibrary{shapes.geometric, arrows}

\definecolor{color1}{HTML}{6bd2db}
\definecolor{color2}{HTML}{0ea7b5}
\definecolor{color3}{HTML}{0c457d}
\definecolor{color4}{HTML}{ffbe4f}
\definecolor{color5}{HTML}{e8702a}

@@

\author{
    \href{https://orcid.org/0009-0001-0613-7978}{
        \includegraphics[scale=0.06]{orcid.pdf}
        \hspace{1mm}Jonas Pleyer
    }
    \thanks{
        \href{https://jonas.pleyer.org}{jonas.pleyer.org},
        \href{https://cellular-raza.com}{cellular-raza.com}
    }\\
	Freiburg Center for Data-Analysis and Modeling\\
	University of Freiburg\\
	\texttt{jonas.pleyer@fdm.uni-freiburg.de} \\
	%% examples of more authors
	\And
	\href{https://orcid.org/0000-0002-6371-4495}{
        \includegraphics[scale=0.06]{orcid.pdf}
        \hspace{1mm}Christian Fleck
    }\\
	Freiburg Center for Data-Analysis and Modeling\\
	University of Freiburg
}

@@

\begin{figure}[H]
    \centering
    \tikzstyle{rect} = [
        rectangle,
        % rounded corners,
        minimum width=4cm,
        minimum height=1cm,
        text centered,
        fill=color4!70,
    ]
    \tikzstyle{brect} = [rect, fill=color5!50]
    \tikzstyle{frect} = [rect, fill=color3!40]
    \tikzstyle{arrow} = [thick,->,>=stealth]
    \tikzstyle{darrow} = [arrow, dashed]
    \begin{tikzpicture}
        \node (model) [brect, align=center] at (0,0) {Mathematical \&\\ Computational Model};
        \node (extraction) [rect] at (0, -2) {Data Extraction};
        \node (params) [brect] at (6, -2) {Parameter Estimation};
        \node (visualization) [rect] at (6, 0) {Visualization};

        \draw[arrow] (model) -- (extraction);
        \draw[darrow] (params) -- node[anchor=south, rotate=-18.43] {validate} (model);
        \draw[arrow] (model) -- (visualization);
        \draw[arrow] (visualization) -- (params);
        \draw[arrow] (extraction) -- (params);

        % Future stuff
        \node (realimages) [frect] at (12, 0) {Data Generation};
        \node (segmentation) [frect] at (12, -2) {Cell Segmentation};
        \node (tracking) [frect] at (12, -3) {Cell Tracking};
        \node (analysis) [frect] at (0, -4) {Analysis Methods};

        \draw[arrow] (visualization) -- (realimages);
        \draw[arrow] (realimages) -- (segmentation);
        \draw[arrow] (extraction) -- (analysis);
        \draw[arrow] (params) -- (realimages);
    \end{tikzpicture}
    \caption{
        The project can be divided into 4 subparts (orange + yellow) that interact with and depend
        on each other.
        The orange blocks require input by the user by specifying the model and supplying input
        data.
        Additional blocks (blue-gray) indicate possible future extensions of this project.
        Arrows with solid lines indicate dependencies between the subparts while dashed arrows carry
        special meaning, which is specifically annotated.
        The mathematical models provides the theoretical framework and is thus the parent of all
        following work.
        "Data Extraction" allows us to read variable and parameter values directly from microscopic
        images.
        The obtained data can then be used to initialize a numerical simulations and thus estimate
        parameters.
        On the other hand, results produced by our model can be visualized as 3D renders and
        projected to cell-masks.
        These algorithms can be exploited to construct cost-functions which can be used in the
        parameter estimation process.
        In future work, this visualization scheme could be extended to generate realistic
        microscopic images with cell-masks and thus train cell-segmentation and cell-tracking
        algorithms.
    }
    \label{fig:flowchart-project-structure}
\end{figure}

@@

@@

% --------------------------------------------------------------------------------------------------
\subsection{Data Generation}

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[hidelinks]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,amssymb,amsthm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{float}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{mathrsfs}
\usepackage{onimage}
\usepackage{graphicx}
\usepackage{doi}
\usepackage{acronym}
\usepackage{listings}
\usepackage[mode=buildnew]{standalone}
\usepackage{tikz}
\usepackage{tabularx}
\usepackage{siunitx}
\usepackage{cite}
\usepackage{acronym}
\usepackage{xcolor}

\input{figures/tikz-preamble.tex}

@@

\author{
    \href{https://orcid.org/0009-0001-0613-7978}{
        \includegraphics[scale=0.06]{figures/orcid.pdf}
        \hspace{1mm}Jonas Pleyer
    }
    \thanks{
        \href{https://jonas.pleyer.org}{jonas.pleyer.org},
        \href{https://cellular-raza.com}{cellular-raza.com}
    }\\
	Freiburg Center for Data-Analysis and Modeling\\
	University of Freiburg\\
	\texttt{jonas.pleyer@fdm.uni-freiburg.de} \\
	%% examples of more authors
	\And
	\href{https://orcid.org/0000-0002-6371-4495}{
        \includegraphics[scale=0.06]{figures/orcid.pdf}
        \hspace{1mm}Christian Fleck
    }\\
	Freiburg Center for Data-Analysis and Modeling\\
	University of Freiburg
}

@@

\begin{figure}[H]
    \centering
    % \includegraphics[width=0.8\textwidth]{overview.pdf}
    \includestandalone[width=0.8\textwidth]{figures/concept/overview}
    \caption{
        The project can be divided into 4 subparts (yellow) that interact with and depend on each
        other.
        Additional blocks (blue-gray) indicate possible future extensions of this project.
        Arrows with solid lines indicate dependencies between the subparts while dashed arrows carry
        special meaning, which is specifically annotated.
        The mathematical models provides the theoretical framework and is thus the parent of all
        following work.
        "Data Extraction" allows us to read variable and parameter values directly from microscopic
        images.
        The obtained data can then be used to initialize a numerical simulations and thus estimate
        parameters.
        On the other hand, results produced by our model can be visualized as 3D renders and
        projected to cell-masks.
        These algorithms can be exploited to construct cost-functions which can be used in the
        parameter estimation process.
        In future work, this visualization scheme could be extended to generate realistic
        microscopic images with cell-masks and thus train cell-segmentation and cell-tracking
        algorithms.
    }
    \label{fig:flowchart-project-structure}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_fit/snapshots/exact-004444.png}%
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.33\textwidth]
        {figures/crm_fit/snapshots/exact-004444.png}%
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
\end{figure}

@@
